### 2019-06-29 Memoria/Chapters/ExperimentalResults.tex
Escrito analisis de resultados en memoria

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Restarts per generation with $\xi \le 0$]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMenor.pdf}} 
	\hspace{1cm}
	\subfloat[Restarts per generation with $\xi \ge 0$]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMayor.pdf}}
	\caption{Visual representation of the influence of $\xi$ over the \acs{DILS}$_{CC}$ optimization process.}
	\label{fig:RestartsLines}
\end{figure}

@@

We can conclude that the parameter $\xi$ has the expected influence over the \acs{DILS}$_{CC}$ optimization process. Note that in a problem in which the \acs{LS} procedure used is deterministic we would also find differences between consecutive runs over the same datasets, given that the \acs{DILS} general optimization scheme uses stochastic tools---such as the recombination and mutation operators---to make sure that diversity is introduced in the search of the global optima for the problem we apply it to.

@@

Regarding the comparison between the two new proposals---\acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$---, we have seen how both of them represent an improvement over previous constrained clustering approaches, being \acs{DILS}$_{CC}$ the best of the two. This is due to the fact that a real-based problem representation does not provide any other advantage that the possibility to apply real-coding evolutionary methods. This way, given that the \acs{SHADE} method proved to be the best approach to real optimization problems we expected it to outperform \acs{BRKGA}, though \acs{SHADE} imposes an arithmetic relationship between individuals---through the mutation operator---which is hard to understand in the clustering domain. This is, applying an arithmetic operator to two labels does not provide easily understandable combination of them. Judging from the raw results, and with its straight integer-based problem representation and its strong exploitation capability \acs{DILS}$_{CC}$ stands out as the best of all methods we obtained results for.

>>>>

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Restarts per generation with $\xi \le 0$]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMenor.pdf}} 
	\hspace{1cm}
	\subfloat[Restarts per generation with $\xi \ge 0$]
	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMayor.pdf}}
	\caption[Visual representation of the influence of $\xi$ over the DILS$_{CC}$ optimization process.]{Visual representation of the influence of $\xi$ over the \acs{DILS}$_{CC}$ optimization process.}
	\label{fig:RestartsLines}
\end{figure}

@@

We can conclude that parameter $\xi$ has the expected influence over the \acs{DILS}$_{CC}$ optimization process. Note that in a problem in which the \acs{LS} procedure used is deterministic we would also find differences between consecutive runs over the same datasets, given that the \acs{DILS} general optimization scheme uses stochastic tools---such as the recombination and mutation operators---to make sure that diversity is introduced in the search of the global optima for the problem we apply it to.

@@

All comparisons carried out in this sections are merely supported by the raw results. To prove whether the new proposals represent an improvement over the state-of-the-art methods it is necessary to perform an statistic-based empirical analysis of the results. The last chapter (Chapter \ref{ch:AnalysisResults}) of Part \ref{pt:Experiments} will address this question.

### 2019-06-29 Memoria/Chapters/Introduction.tex
Escrito analisis de resultados en memoria

Among the four main categories of machine learning methods we find the \acf{SSL}, which is the focus of this work. As we have already mentioned, it is a machine learning paradigm that arises from adding incomplete information to unsupervised learning. We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}. The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances. The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this work \cite{triguero2015self}.

\section{Objetivos}

\section{Estructura}

>>>>

Among the four main categories of machine learning methods we find the \acf{SSL}, which is the focus of this work. As we have already mentioned, it is a machine learning paradigm that arises from adding incomplete information to unsupervised learning. We can divide \acs{SSL} methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}. The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances. The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this work \cite{triguero2015self}.

\section{Structure}

### 2019-06-29 Memoria/Chapters/NewProposals.tex
Correcciones y nuevas secciones
Escrito analisis de resultados en memoria

Finding the optimal partition in a dataset, with respect to any kind of reasonable criteria, is known to be a $\mathbf{NP}$-hard problem. As we have already mentioned in Section \ref{sec:FeasibilityProblem}, the incorporation of constraints may modify the complexity of the clustering problem, making it $\mathbf{NP}$-complete and therefore intractable \cite{davidson2005clustering}. Bearing this in mind, in this chapter we propose two novel approaches to find approximate solutions to the constrained clustering problem. In Section \ref{sec:CCSHADE} a \acf{DE} based approach is presented, while in Section \ref{sec:DILS} a new \acf{LS} variant is discussed, as well as its application to constrained clustering.

@@

\section{Brief Review of the BRKGA Algorithm} \label{sec:brkga}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.4]{gfx/NewProp/SHADE/BRKGA_Diagram.jpg} 
	\caption[BRKGA$_{CC}$ optimization process for constrained clustering.]{\acs{BRKGA}$_{CC}$ optimization process for constrained clustering. \cite{de2017comparison}.}\label{fig:BRKGA_CC}
\end{figure}

@@

\section{SHADE optimization scheme for constrained clustering} \label{sec:SHADEadapt}

@@

\begin{figure}[bth]
	\myfloatalign
	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/Dataset}} \quad
	%\subfloat[Imagen original]
	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C0}} \quad
	%\subfloat[Clustering sin restricciones]
	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C1}} \quad
	%\subfloat[Clustering con restricciones]
	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C2}} \quad
	\caption[Three partitions over a toy dataset.]{Three partitions over a toy dataset.}
	\label{fig:toydatasets}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.4]{gfx/NewProp/SHADE/SHADE_Diagram.jpg} 
	\caption[SHADE$_{CC}$ optimization process for constrained clustering.]{\acs{SHADE}$_{CC}$ optimization process for constrained clustering.}\label{fig:SHADE_CC}
\end{figure}

@@

	\caption{Local Search}\label{alg:SHADE_LS}
\end{algorithm}

@@

\subsection{Short Introduction to Iterated Local Search}

@@

\section{The Dual Iterative Local Search Method}

The \acf{DILS} is a new variant of the classic \acs{ILS} method. Its goal is to perform a search in the solution space to find the solution with the best fitness value (given by fitness function $f$) by introducing diversity in an adaptive and guided way. While \acs{ILS} works with a single individual, \acs{DILS} keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima. The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value.

In each generation $G$ of the optimization process \acs{DILS} builds a new individual by applying a recombination operator to $m_b$ and $m_w$. After that, it applies a strong mutation operator to the newly generated individual. This is the individual \acs{DILS} applies the \acs{LS} procedure to in order to improve its fitness value, resulting in the trial individual $m_t$. The trial individual $m_t$ replaces the worst of its parents, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for \acs{DILS}. Equation \ref{eq2.1} displays the expression for this criterion in the case of a minimization problem.

@@

\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method. This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best parent. Figure \ref{img:DILS_diag} summarizes the \acs{DILS} optimization process.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.25]{gfx/NewProp/DILS/DILSDiagram.jpg}
	\caption{Diagram summarizing \acs{DILS} optimization process.}\label{img:DILS_diag}
\end{figure}

@@

\section{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}

In this section, we present an optimization scheme for the \acs{DILS} algorithm when applied to constrained clustering. We will call this new approach \acs{DILS}$_{CC}$. We will use an integer-based representation for all the individuals \acs{DILS}$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---with $n$ being the number of instances of the dataset. Each integer $m_i \in [0,k)$ in $m$ represents the label of the cluster that instance $x_i$ is assigned to, and therefore each individual $m$ is a solution to the clustering problem. Additionally, in order not to bias the exploration of the solution space, we will use random initialization to set $m_b$ and $m_w$; this procedure will also be used in the reinitialization of $m_w$.

@@

	\caption{Local Search}\label{alg:DISL_LS}
\end{algorithm}

>>>>

Finding the optimal partition in a dataset, with respect to any kind of reasonable criteria, is known to be a $\mathbf{NP}$-hard problem. As we have already mentioned in Section \ref{sec:FeasibilityProblem}, the incorporation of constraints may modify the complexity of the clustering problem, making it $\mathbf{NP}$-complete and therefore intractable \cite{davidson2005clustering}. Bearing this in mind, in this chapter we propose two new approaches to find approximate solutions to the constrained clustering problem. In Section \ref{sec:CCSHADE} a \acf{DE} based approach is presented, while in Section \ref{sec:DILS} a new \acf{LS} variant is discussed, as well as its application to constrained clustering.

@@

\subsection{Brief Review of the BRKGA Algorithm} \label{sec:brkga}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.31]{gfx/NewProp/SHADE/BRKGA_Diagram.jpg} 
	\caption[BRKGA$_{CC}$ optimization process for constrained clustering.]{\acs{BRKGA}$_{CC}$ optimization process for constrained clustering. \cite{de2017comparison}.}\label{fig:BRKGA_CC}
\end{figure}

@@

\subsection{SHADE optimization scheme for constrained clustering} \label{sec:SHADEadapt}

@@

\begin{figure}[bth]
	\myfloatalign
	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/Dataset}} \quad
	%\subfloat[Imagen original]
	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/C0}} \quad
	%\subfloat[Clustering sin restricciones]
	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/C1}} \quad
	%\subfloat[Clustering con restricciones]
	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/C2}} \quad
	\caption[Three partitions over a toy dataset.]{Three partitions over a toy dataset.}
	\label{fig:toydatasets}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.31]{gfx/NewProp/SHADE/SHADE_Diagram.jpg} 
	\caption[SHADE$_{CC}$ optimization process for constrained clustering.]{\acs{SHADE}$_{CC}$ optimization process for constrained clustering.}\label{fig:SHADE_CC}
\end{figure}

@@

	\caption[SHADE$_{CC}$ Local Search]{\acs{SHADE}$_{CC}$ Local Search}\label{alg:SHADE_LS}
\end{algorithm}

@@

\subsection{A Glimpse to Iterated Local Search}

@@

\subsection{The Dual Iterative Local Search Method}

The \acf{DILS} is a new variant of the classic \acs{ILS} method. Its goal is to perform a search in the solution space to find the solution with the best fitness value (given by fitness function $f$) by introducing diversity in an adaptive and guided way. While \acs{ILS} works with a single individual, \acs{DILS} keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima. The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value at the end of each stage (generation) of the optimization process.

In each generation $G$ of the optimization process \acs{DILS} builds a new individual by applying a recombination operator to $m_b$ and $m_w$. After that, it applies a strong mutation operator to the newly generated individual. This is the individual \acs{DILS} applies the \acs{LS} procedure to in order to improve its fitness value, resulting in the trial individual $m_t$. The trial individual $m_t$ replaces the worst of its predecessors, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for \acs{DILS}. Equation \ref{eq2.1} displays the expression for this criterion in the case of a minimization problem.

@@

\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method. This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best predecessor. Figure \ref{img:DILS_diag} summarizes the \acs{DILS} optimization process.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.27]{gfx/NewProp/DILS/DILSDiagram.jpg}
	\caption{Diagram summarizing \acs{DILS} optimization process.}\label{img:DILS_diag}
\end{figure}

@@

\subsection{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}

In this section, we present the application scheme of \acs{DILS} to constrained clustering. We will call this new approach \acs{DILS}$_{CC}$. We will use an integer-based representation for all the individuals \acs{DILS}$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---with $n$ being the number of instances of the dataset. Each integer $m_i \in [0,k)$ in $m$ represents the label of the cluster that instance $x_i$ is assigned to, and therefore each individual $m$ is a solution to the clustering problem. Additionally, in order not to bias the exploration of the solution space, we will use random initialization to set $m_b$ and $m_w$; this procedure will also be used in the reinitialization of $m_w$.

@@

	\caption[DILS$_{CC}$ Local Search]{\acs{DILS}$_{CC}$ Local Search}\label{alg:DISL_LS}
\end{algorithm}

### 2019-06-29 Memoria/FrontBackmatter/Declaration.tex
Escrito analisis de resultados en memoria

%*******************************************************
% Declaration
%*******************************************************
\refstepcounter{dummy}
\pdfbookmark[0]{Declaration}{declaration}
\chapter*{Declaration}
\thispagestyle{empty}
\textbf{JAVIER LEÓN PALOMARES} con DNI $77138857$P, actuando en su propio nombre y derecho, \\

\textbf{DECLARA, BAJO SU RESPONSABILIDAD:} \\

Que el Trabajo Fin de Máster presentado en la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicación de la Universidad de Granada, con fecha \_\_ de julio de 2019, titulado <<\textit{\_\_}>> es original, no es copia ni adaptación de ningún otro trabajo, inédito, y no ha sido difundido por ningún medio, ni presentado anteriormente por quien suscribe o por otra persona. \\

Y para que así conste a los efectos oportunos, firmo la presente en Granada, a \_\_ de julio de 2019.

@@

%\noindent\textit{\myLocation, \myTime}

@@

\begin{flushright}
    \begin{tabular}{m{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}

Asimismo, yo, Javier León Palomares, alumno del Máster en Ciencia de Datos e Ingeniería de Computadores de la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicación de la Universidad de Granada, con DNI $77138857$P, autorizo la ubicación de la siguiente copia de mi Trabajo Fin de Máster en la biblioteca del centro para que pueda ser consultada por las personas que lo deseen.

\begin{flushright}
    \begin{tabular}{m{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}

>>>>

%*******************************************************
% Declaration
%*******************************************************
\refstepcounter{dummy}
\pdfbookmark[0]{Declaration}{declaration}
\chapter*{Declaración de Originalidad}
\thispagestyle{empty}

@@

\noindent \textbf{D. GERMÁN GONZÁLEZ ALMAGRO} con DNI $76593910$T, actuando en su propio nombre y derecho,\\

\noindent \textbf{DECLARA, BAJO SU RESPONSABILIDAD}:\\

\noindent Que el Trabajo Fin de Grado presentado en la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicaciones de la Universidad de  Granada, con fecha 17 de Junio de 2018, titulado <<\textit{Constrained Clustering: New Proposals}>> es original, no es copia de ningún otro trabajo, inédito, y no ha sido difundido por ningún medio, ni presentado anteriormente por quien suscribe o por otra persona.\\

\noindent Y para que así conste a los efectos oportunos, firmo la presente en Granada, a 17 de junio de 2018.

\vspace{3cm}

@@

\begin{flushright}
    \begin{tabular}{m{6cm}}
        \\ \hline
        \centering\textbf{Fdo. \myName} \\
    \end{tabular}
\end{flushright}

### 2019-06-29 Memoria/FrontBackmatter/Legality.tex
Escrito analisis de resultados en memoria

Yo, \textbf{Germán González Almagro}, alumno de la titulación en ingeniería informática de la \textbf{Escuela Técnica Superior
de Ingenierías Informática y de Telecomunicación de la Universidad de Granada}, con DNI $\mathrm{76593910}$T, autorizo la
ubicación de la siguiente copia de mi Trabajo Fin de Grado en la biblioteca del Centro para que pueda ser
consultada por las personas que lo deseen.

@@

Que el presente trabajo, titulado \textit{\textbf{<<Clustering con Restricciones: un marco unificado>>}},
ha sido realizado bajo su supervisión por \textbf{Germán González Almagro}, y autorizamos la defensa de dicho trabajo ante el tribunal
que corresponda.

@@

Y para que conste, expiden y firman el presente informe en Granada a 17 de junio de 2018.

>>>>

Yo, \textbf{Germán González Almagro}, alumno de la titulación en ingeniería informática de la \textbf{Escuela Técnica Superior
de Ingenierías Informática y de Telecomunicación de la Universidad de Granada}, con DNI $\mathrm{76593910}$T, autorizo la
ubicación de la siguiente copia de mi Trabajo Fin de Máster en la biblioteca del Centro para que pueda ser
consultada por las personas que lo deseen.

@@

Que el presente trabajo, titulado \textit{\textbf{<<Constrained Clustering: New Proposals>>}},
ha sido realizado bajo su supervisión por \textbf{Germán González Almagro}, y autorizo la defensa de dicho trabajo ante el tribunal
que corresponda.

@@

Y para que conste, expido y firmo el presente informe en Granada a 17 de junio de 2018.
